---
title: "Gov 50 Final Project"
author: "Melih Cevik"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---

## Project thoughts


  Can the percentage of a town's fox news subscribers predict their GOP vote share. In this study, I plan to examine whether or not there is a correlation between the percentage of fox news subscribers, and whether the percentage of how many voted Republican in the 2000 presidential election. I hypothesize that towns with a greater percentage of fox news subscribers will see a greater percentage of those voting Republican. My explanatory variable is the share of subscribers to fox news in a town in the year 2000 , my response variable is Republican vote share in that town in the year 2000. If I observe, on average, a greater republican vote percentages in a town where there is a high prevalence of fox news, then my hypothesis will be correct. If I observe on average, a lower republican vote percentage in a town where there is a high share of fox news subscribers. 
  In addition to the linear regression that I will run on the overall dataset, I will also run a linear regression when controlling for college education. Furthermore I will also see if there are variances on whether or not there is a correlation depending on the state in which an individual resides in. I plan on examining if there are differences in correlation between solidly red or blue states and swing states. I find this all interesting because it is an interesting way to determine whether the media changes our opinions or solely informs us.

```{r}
library(tidyverse)
library(infer)
library(broom)
library(modelsummary)
foxnews <- read_csv("foxnews.csv", show_col_types = FALSE)
```





```{r}
foxnews1 <- foxnews |> 
  filter(!subrf2000 == 0.00000000 ) |>
  select(state,town,subrf2000,gopvoteshare1996,gopvoteshare2000 ) 

ggplot(data = foxnews1, mapping = aes(x = subrf2000, y = gopvoteshare2000)) + 
  geom_point(color = "gray", shape = 1) +  
  geom_smooth(method = "lm", se = FALSE) +  
  labs(x = "Share of Households who subscribe to Fox News in the Town (as a percent)",
       y = "Republican vote(%) in Town",
       title = "The relation between Fox News Subscriptions and the 2000 Presidential Election",source = "The Fox News Effect: Media Bias and Voting. Published in the Quarterly Journal of Economics") +
  theme_bw()

```
  As we see in this graph, as the share of households who subscribe to Fox news in the Town as a percentage increases, on average the Republican vote share in the town slightly decreases. For every one percentage point increase in a Town's republican vote share in the 2000 election, the GOP vote share decreased by 0.018 percentage points.This would mean that my hypothesis that there was a positive correlation between Fox News subscriptions and the GOP vote share in the 2000 election is proven false.
  It is important to note that a critical confounding variable exists within the research we conduct and the data set. While the data shows the subscribers as a percentage, it is limited in the fact that we do not know whether or not an individual is a viewer. This is a cross-sectional research project. 



```{r}
var_labels <- c(
"(Intercept)" = "Intercept",
"subrf2000" = "Subscribers to FNC in 2000",
"college1990" = "Percent of Town with College Degree"
)



fit <- lm(gopvoteshare2000 ~ subrf2000, data = foxnews )
fit2 <- lm(gopvoteshare2000 ~ subrf2000 + college1990, data = foxnews )


modelsummary::modelsummary(list (fit,fit2),statistic = c("s.e. = {std.error}",
"p = {p.value}"),
coef_map = var_labels,
gof_map = c("nobs", "r.squared", "adj.r.squared"))




```

  The analysis I decided to conduct was a linear regression output. The output describes the result of both the graph above and when we hold  for the variation as a result of college education, which is graphed below. The dependent variable in both linear regression models is a town's GOP vote share as a percentage in the 2000 presidential election. The independent variable in both linear regression models is the town's subscription rate, measured in a percentage. 
  Our main coefficient of interest in the 1st linear regression model, of -0.019, means that for every percentage point increase in a town of Fox News Subscribers in the year 2000, there was a decrease for Republican support in the 2000 election of 0.019 on average. Our p-value here of 0.024 is less than a 0.05 significance level, as a result we would reject the null hypothesis. 
  On a secondary analysis which is graphed below, we controlled for college education level of that town, what proportion had a college education in 1990. Our main coefficient of interest in the 2nd linear regression model was -0.01. This means that for every percentage point increase in a town of Fox News Subscribers in the year 2000, there was a decrease for Republican support in the 2000 election of 0.01 on average. In this regression our p-value of 0.246 is greater than a 0.05 significance level, therefore we would fail to reject the null hypothesis. 

```{r}
ggplot( data = fit2, mapping = aes(x = subrf2000, y = gopvoteshare2000)) + 
  geom_point(color = "gray", shape = 1)+  
  geom_smooth(method = "lm", se = FALSE) +  
  labs(x = "Share of Households who subscribe to Fox News in the Town (as a percent)",
       y = "Republican vote(%) in Town",
       title = "Fox News Subscriptions and the 2000 Election controlled for College Education",
       source = "The Fox News Effect: Media Bias and Voting. Published in the Quarterly Journal of Economics") +   theme_bw()
```
As we see in this graph, as the share of households who subscribe to Fox news in the Town as a percentage increases, on average the Republican vote share in the town slightly decreases. For every one percentage point increase in a Town's republican vote share in the 2000 election, the GOP vote share decreased by 0.01 percentage point when we were holding for college education level. This shows that college education levels to some extent were confounding variables in determining the impact of Fox News subscriptions on a towns Republican vote share.
  As seen in our main analysis, it is important to note that a critical confounding variable exists within the research we conduct and the data set. While the data shows the subscribers as a percentage, it is limited in the fact that we do not know whether or not an individual is a viewer. 




```{r}
foxnews2 <- foxnews |> 
select(state,town,subrf2000,gopvoteshare1996,gopvoteshare2000 ) 
foxnews2 <- foxnews2 |>
filter(!subrf2000 == 0.00000000 ) |>
filter(state ==  c("Ca", "Al", "Pa"))

ggplot(data = foxnews2, mapping = aes(x = subrf2000, y = gopvoteshare2000)) +
  geom_point(color = "gray", shape = 1) +  
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ state) + 
  labs(x = "Share of Households who subscribe to Fox News in the Town (as a percent)",
       y = "Republican vote(%) in Town",
       title = "Fox News Subscriptions and Republican vote % in the 2000 Election",
       source = "The Fox News Effect: Media Bias and Voting. Published in the Quarterly Journal of Economics")
```
  A question that arises is whether or not Fox News subscriptions had different effects on Vote share's depending on a location. To determine this, we selected three states, one that solidly voted Republican (Alabama), one that solidly voted Democratic (California) and a state that was decided within 5% (Pennsylvania). Our analysis shows that Fox News Subscriptions had different correlations in different states. In the noncompetitive states, there was a positive correlation between Fox News Subscriptions and GOP vote share.However in Pennsylvania the swing state, there was a negative correlation. 





Conclusion: 
  Overall, we find that in the 2000 Presidential election, there was a slightly negative correlation between a town's percentage of Republican vote share in the 2000 Presidential election, and the percentage of that town who has a cable subscription that includes Fox News. When controlling for college education, we see that this negative correlation still exists, despite a weakening. We also find that there is variation in correlation dependent on the individual state. This is the opposite of my original hypothesis, however, there were numerous limitations and confounding variables to allow us to make an inference. First, the dataset is outdated, there have since been 5 presidential election cycles, as a result we cannot apply the conclusion to today's political situation. Furthermore, a confounding variable that existed within the data was whether or not the town watched Fox News. If I were to design a study today with unlimited time and money, I would look at in the next presidential election, a town's viewership, not subscription to Fox News and their Republican vote share. I would also collect more data points, this data set failed to collect data from a wide variety of states and localities, I would have a broader sample if I had unlimited resources. 




