---
title: "Gov 50 Final Project"
author: "Melih Cevik"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---

## Project thoughts


Can the percentage of a town's fox news subscribers predict their GOP vote share. In this study, I plan to examine whether or not there is a correlation between the percentage of fox news subscribers, and whether the percentage of how many voted Republican in the 2000 presidential election. I hypothesize that towns with a greater percentage of fox news subscribers will see a greater percentage of those voting Republican. My explanatory variable is the share of subscribers to fox news in a town in the year 2000 , my response variable is Republican vote share in that town in the year 2000. If I observe, on average, a greater republican vote percentages in a town where there is a high prevalence of fox news, then my hypothesis will be correct. If I observe on average, a lower republican vote percentage in a town where there is a high share of fox news subscribers.


```{r}
library(tidyverse)
library(infer)
library(broom)
foxnews <- read_csv("foxnews.csv")
```





```{r}
foxnews1 <- foxnews |> 
  filter(!subrf2000 == 0.00000000 ) |>
  select(state,town,subrf2000,gopvoteshare1996,gopvoteshare2000 ) 

ggplot(data = foxnews1, mapping = aes(x = subrf2000, y = gopvoteshare2000)) + geom_point() +  geom_smooth(method = "lm", se = FALSE) +  labs(x = "Share of Households who subscribe to Fox News in the Town (as a percent)",
       y = "Proportion of who Voted Republican in the Town in the 2000 election",
       title = "The relation between Fox News Subscriptions and the 2000 Presidential Election",
       source = "The Fox News Effect: Media Bias and Voting. Published in the Quarterly Journal of Economics")

```

```{r}
fit <- lm(gopvoteshare2000 ~ subrf2000, data = foxnews1 )
fit
```


